Composition-climate paper

### 2015-07-09 Composition_Ecology.Rmd
Added some references, and more extensive text.

Over the past 200 years centennial scale climate variability and land use change have interacted to transform the forests of the northeastern United States. There are clear indications that modern forests, as represented by the Forest and Inventory Analysis show greater homogeneity and significantly different species associations than pre-settlement forests [@goring2015preset;@schulte2007homogenization]. This shift has occured over a period during which climate has begun to show the effects of anthropogenic warming, and a period in which we have increasingly used the relationship between modern tree species distributions and climate to understand future responses of tree species and forest assemblages to climate change [@iverson2013tree;@iverson1998predicting].  One of the most significant changes to forest tree distributions in the upper midwest is the loss of extensive cover through land use change [@rhemtulla2009legacies], particularly the .

@@

>>>>

Over the past 200 years centennial scale climate variability and land use change have interacted to transform the forests of the northeastern United States. There are clear indications that modern forests, as represented by the Forest and Inventory Analysis show greater homogeneity and significantly different species associations than pre-settlement forests [@goring2015preset;@schulte2007homogenization]. This shift has occured over a period during which climate has begun to show the effects of anthropogenic warming, and a period in which we have increasingly used the relationship between modern tree species distributions and climate to understand future responses of tree species and forest assemblages to climate change [@iverson2013tree;@iverson1998predicting].  One of the most significant changes to forest tree distributions in the upper midwest is the loss of extensive cover through land use change [@rhemtulla2009legacies], particularly along the historical prairie and savanna margins of the region (Figure 1), where extensive land use conversion has left little natural land.

@@

The interaction of land use change and climatic change has been of interest to conservation managers and scientists [e.g., @pyke2004habitat] since managed natural areas often represent a biased selection from the available environmental space [@scott2001nature].  Pyke [-@pyke2004habitat] lays out a useful framework for assessing the interaction of climate and land use change.  By deliniating three tpyes of interaction, compunding, confounding and contradicting, it is possible to understand both the nature of the interaction and the potential effect on species resiliance and response to broader regional change.

Compounding effects occur when both climatic change and land use change are biased in the same direction, *e.g.*, warming climate is compounded by land use change that is biased towards warmer environments.  In the case of warming a species may lose climatic suitability along the warmest margins of its range at the same time as land use pressure increases, hastening decline in response to warming.  Confounding effects may cause little overall effect on species distributions since in this case the land use pressure falls within the central distribution for the environmental variable, they may increase the variability of predictive models, but do little to the overall envelope.  Contradicting effects result in a narrower climate niche for taxa, where warming forces the distribution mean higher as populations in the warmer tail of the distribution experience greater environmental pressure, and, secondarily environmentally biased land use pressure in the cooler end of the tail excludes the taxa from those regions.
